---
title: "Rmarkdown is pretty dang rad" # a fun title!
author: "Nicolita the Snail" # a fun name!
date: "`r format(Sys.time(), '%d %B %Y - %H:%M')`" # automatically inputs the date and time - neato
output: 
  html_document: # for word doc output, change this to word_document
                 # for pdf output, change this to pdf_output
    toc: true # table of contents
    toc_float: false #floats at the side of the html
    toc_collapsed: false
    toc_depth: 6 # can have up to 6 levels in the table of contents
    number_sections: true # numbers the sections
theme: lumen # a nice crisp theme
---

This is the basic rmakrdown template I use each time. The html output is nice and interactive, and the pdf is clean. Remember to write 2 spaces after a paragraph or sentance to start a new line.    

If you want more whitespace between paragraphs or figures in the html output, use this code between sections:  

<div style="margin-bottom:50px;">
</div>

# This is a big ole heading!
## This is a slightly smaller heading!
### Even smaller heading!
#### Does it keep going?
##### OMG how far does it go?
###### Its headings all the way down!!!

<div style="margin-bottom:50px;">
</div>

```{r setup, echo = FALSE}
# sets default to echo= TRUE, which will show your code chunks in the html output
# as a bit of irony, putting echo = FALSE in the curly brackets above will hide this code chunk from the html output. Do this for any coding that you don't need to show your audience. 
knitr::opts_chunk$set(echo = TRUE)
```

```{r clean space, echo = FALSE}
# always start off with a clean slate....then mess it up
# echo = FALSE hides the code
rm(list = ls())
```

```{r load packages, message = FALSE,  echo = FALSE, warning = FALSE}
library(dplyr) # or the tidyverse package
library(colorspace) # i love this package for vision-diverse friendly colour palletes
library(ggplot2)
library(ggpubr) #for multiplots
library(gtsummary) # for summary tables
library(knitr) # for rmarkdown
library(kableExtra) # for basic tables

# warning = FALSE hides the warnings from the html output
```
<div style="margin-bottom:50px;">
</div>

## Data
```{r datasets, warning = FALSE}
df <- read.csv("salpref_smoltdata_20102020.csv")
# Rmarkdown is handy to annotate any changes you make to datasets. I like it because a broken script will not knit so youre forced to deal with problems as they arrise. 

# rename columns to be more succinct and coding friendly
df <- df %>%
  dplyr::select(-X) %>% # remove unneccessary columns
                        # note, if you get an error with a pipe in dplyr, sometimes its due to conflict with plyr, so you have to either unload plyr or specify that the package you want is dplyr
  rename(fish.id = Fish_ID,
         experiment = Experiment,# rename works as NEW = OLD
         outmigr.wk = OutMigration_Week, 
         outmigr.day = Days_Since_Outmigr, 
         fl = Field_FL_mm, # also make sure you have documentation of your units somewhere (like here!) if you remove them from your column names
         mass = Field_Weight_g, 
         k = Fulton_K, 
         explore = Active, # exploratory = 1, non-exploratory = 0
         pref = Preferred_Chamber_p0.5) # salinity of preferred chamber

attach(df)
summary(df)
df <- df %>% 
  mutate_at(vars(outmigr.wk, explore), funs(factor))

# order salinity preference
df$pref <- factor(df$pref, levels = c("Freshwater", "Brackish", "Saltwater"))
```
<div style="margin-bottom:50px;">
</div>

## Figures
```{r Figures, fig.cap= "Figure 1. Fish Condition by Outmigration Week ", echo = FALSE, message = FALSE, include = TRUE, out.width= '50%',  fig.show ='hold', warning = FALSE}
# fig.cap adds a figure caption just to the html file after you knit it
# out.width scales your figures, if you are doing a multipanel just to present in the html, make it 50% so that its 2 columns. This wont save the figure, just show it. To save a multiplot, use the code below

margins <- c(0.5,0.5,0.5,1.1) # set ur margins to publ. standard. This is a good staple. 

f1 <- ggplot(df, aes(outmigr.wk, k, fill = pref))+
  geom_point(position=position_dodge(width=0.75), aes(group=pref))+
  geom_boxplot()+
  scale_fill_discrete_sequential(palette = "BluGrn") +
  labs(title = "Composite",
       fill = "Preference",
       y = expression(bold(Fulton~K~"("~g~"*"~mm^{"-3"}~")")), 
       x = "\n Outmigration Week") + # the 
  scale_x_discrete(labels = c('River','Estuary','Ocean')) +
  theme_classic() +
  theme(title = element_text(size=14, face = "bold"), 
        axis.text=element_text(size=12, colour="black"),
        legend.position = c(0.9, 0.97),
        axis.title=element_text(size=14, face = "bold"), 
        legend.title =element_text(size=9, colour="black"),
        legend.text =element_text(size=7, colour="black"),
        plot.margin=unit(margins,"line"), 
        strip.text=element_text(hjust=0), 
        panel.background = element_rect(fill = "transparent"), #transparent background baaabyyy
        plot.background = element_rect(fill = "transparent", color = NA), 
        axis.text.x = element_text()) 

#show it
print(f1) 

#save it
ggsave("yourfirstfig.jpeg",plot=f1,units="in",height=6,width=8,dpi=800) # you can change the output size and dpi here. This is a good staple. 

# no trend as aggregate data, but we have included both exploratory and non exploratory fish, so lets filter the dataset by that

df_nonexpl <- df %>%
  filter(explore == '0')

f2 <- ggplot(df_nonexpl, aes(outmigr.wk, k, fill = pref))+
  geom_point(position=position_dodge(width=0.75), aes(group=pref))+
  geom_boxplot()+
  scale_fill_discrete_sequential(palette = "BluGrn") +
  labs(title = "Non-exploratory Fish",
       y = expression(bold(Fulton~K~"("~g~"*"~mm^{"-3"}~")")), 
       x = "\n Outmigration Week") + # the 
  scale_x_discrete(labels = c('River','Estuary','Ocean')) +
  theme_classic()+
  theme(title = element_text(size=14, face = "bold"), 
        axis.text=element_text(size=12, colour="black"),
        legend.position = "none",
        axis.title=element_text(size=14, face = "bold"), 
        plot.margin=unit(margins,"line"), 
        strip.text=element_text(hjust=0), 
        panel.background = element_rect(fill = "transparent"), 
        plot.background = element_rect(fill = "transparent", color = NA), 
        axis.text.x = element_text()) 
print(f2)

df_expl <- df %>%
  filter(explore == '1')

f3 <- ggplot(df_expl, aes(outmigr.wk, k, fill = pref))+
  geom_point(position=position_dodge(width=0.75), aes(group=pref))+
  geom_boxplot()+
  scale_fill_discrete_sequential(palette = "BluGrn") +
  labs(title = "Exploratory Fish",
       y = expression(bold(Fulton~K~"("~g~"*"~mm^{"-3"}~")")), 
       x = "\n Outmigration Week") + # the 
  scale_x_discrete(labels = c('River','Estuary','Ocean')) +
  theme_classic()+
 theme(title = element_text(size=14, face = "bold"), 
        axis.text=element_text(size=12, colour="black"),
        legend.position = "none",
        axis.title=element_text(size=14, face = "bold"), 
        plot.margin=unit(margins,"line"), 
        strip.text=element_text(hjust=0), 
        panel.background = element_rect(fill = "transparent"), 
        plot.background = element_rect(fill = "transparent", color = NA), 
        axis.text.x = element_text()) 
print(f3)

#for a multiplot, I use ggarrange, and then save that new figure. I always have to play around with font sizes to make sure they dont overlap. 
f4 <- ggarrange(f2, f3, f1, ncol=3,nrow=1) #specify number of rows and colums

ggsave("multiplot.jpeg",plot=f4,units="in",height=4,width=13,dpi=800)
```
<div style="margin-bottom:50px;">
</div>

## Tables

<div style="margin-bottom:50px;">
</div>

### Basic
```{r Tables, echo = FALSE, message = FALSE, include = TRUE, results = 'asis'}
#make sure results = 'asis' is there

# use dplyr to group by week, spread by fish behaviour (exploratory or not), and count the number of active fish in each week
df$explore <- dplyr::recode(df$explore, '0'="Non-exploratory", '1' ="Exploratory")
df$outmigr.wk <- dplyr::recode(df$outmigr.wk ,'1'="River", '2' ="Estuary",'4' = "Ocean") # to change grouping name

tab1 <-
  df %>% 
  dplyr::group_by(outmigr.wk, explore) %>%
  dplyr::count(outmigr.wk, explore) %>%
  tidyr::spread(key = explore, value = n ) %>%
  dplyr::rename("Outmigration Week" = outmigr.wk)

#make kable table - output to html, and you can copy paste it directly into word easy peasy
kable(tab1) %>% 
  kable_styling(bootstrap_options = c("striped", "bordered","hover", "condensed", "responsive"), position = "left", full_width = FALSE, fixed_thead = TRUE) # just some nice formatting, you can play around with these

```
<div style="margin-bottom:50px;">
</div>

### Summary
```{r Summary Stats Tables, echo = FALSE, message = FALSE, include = TRUE, results = 'asis'}
#make sure results = 'asis' is there

# select the variables from your dataset that you want to include in your table
tab2 <- df %>% dplyr::select(outmigr.wk, fl, mass, k)


tab2 %>% # your dataset
  #build base summary table 
  tbl_summary(
    #split table by outmigration week
    by=outmigr.wk, 
    #change variable names
    label = list (fl ~ "Fork length (mm)",#now you just write all the variables you want to include and can rename them with units
                  mass ~ "Wet Mass (g)",
                  k ~ "Fulton's K (g/mm^3)"), 
    
    statistic = list(all_continuous() ~ "{mean} ({sd})")) %>% #you can pick the summary stats you want to display
  add_overall() %>%
  add_p(test = list(all_continuous() ~ "aov")) %>% # you can show simple comparative stats here, here one way anova
  as_kable()  %>% 
  kable_styling(bootstrap_options = c("striped", "bordered","hover", "condensed", "responsive"), position = "left", full_width = FALSE, fixed_thead = TRUE) # just some nice formatting, you can play around with these

```